---
title: "Metadata"
author: "Stevens"
date: "2025-02-17"
output: html_document
---

Class exercise to import your project data into R
--Create a folder called "project" in your code folder
--Save this file to the "project" folder
--Move your Nexis spreadsheet to the folder
--Create a project file: File | New Project | *** select Existing folder | "project"
--After project.Rproj launches, then launch this file, metadata.Rmd, and complete the tutorial

**By doing this, you avoid the ring of hell where R can't find your data**


#load tidyverse, janitor, rio
```{r}

library(tidyverse)
library(janitor)
library(rio)
```

```{r}
#Import your data 
bmore <- rio::import("Results list for__crime_ _Baltimore_.XLSX")
bmore1 <- rio::import("Results list for__crime_ _Baltimore_(2).XLSX")
#Combining data files since the export gave me 2 files 
bmore_final <- rbind(bmore, bmore1)

view(bmore_final)

```

```{r}
#Cleaning names 
bmore_final <- bmore_final |> 
  clean_names()

```



```{r}
# fix date
bmore_final <- bmore_final |> 
  mutate(date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))

```



```{r}
#--Using code, describe the number of rows and columns in the dataset. Use the print(Paste0....) to display the results in a sentence form. 
print(paste0("This dataset has ", nrow(bmore_final), " rows and ", ncol(bmore_final), " columns."))
```

```{r}
#--Create a table that displays a count of the top 20 newspaper_city entries. Filter out the NA values
#table for publication location
top <- bmore_final |> 
  #filter nas
  filter(!is.na(publication_location)) |> 
  count(publication_location) |>
  top_n(20, n) |> 
  arrange(desc(n))
top
#but let's see what news organizations these actually are 
top2 <- bmore_final |> 
  #filter nas
  filter(!is.na(publication_4)) |> 
  count(publication_4) |>
  top_n(20, n) |> 
  arrange(desc(n))
top2

top3 <- bmore_final |> 
  #filter nas
  filter(!is.na(publication_16)) |> 
  count(publication_16) |>
  top_n(20, n) |> 
  arrange(desc(n))
top3

```




```{r}
#--Create a table that counts all entries by year.
#create the var year first
bmore_final <- bmore_final |> 
  mutate(year = format(date, "%Y"))


#then do the counts
year <- bmore_final |> 
  count(year) |> 
  arrange(year)

year
```



```{r}
#--Create a simple column ggplot chart that shows the total entries by year
bmore_final |> 
  count(year) |> 
   ggplot(aes(x = year, y = n,
             fill = n)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
#This is your title sequence
  labs(title = "Count of entries by year",
       subtitle = "News Articles on Baltimore Crime",
       caption = "Graphic by Breana Stevens 2-13-14",
       y="Count of entries",
       x="Year")

```


```{r}
#Armed with your data analysis, write a 250 word memo at the end of the R markdown document that summarizes your findings. Use this opportunity to ask questions or note any major flaws in the data.

#I was shocked to see that 27 articles are from international news sources. I didn't realize that such localized crime would get covered on a larger scale. The Baltimore Sun published the most articles in my data set. I was confused why there are two publication data fields (publication_4 and publication_16.. I created tables for both of them and they looked very similar besides the formatting of "The Daily Record". The 100 articles I pulled are from years 2024 and 2025, with slightly more articles coming from 2024. I am excited to dig into keywords that are mentioned in them! 
```

```{r}
#--Compress your folder as a .zip file and upload that to Elms
#okay!
```


# My claude.ai query and reply:
Q: In r, i need to create a new properly formatted date column from a "published_date" column with character data that looks like this: "January 14, 2025 Tuesday",  - how do I transform it into a date format?

A: You can use the as.Date() function along with strptime() to convert that date format. Here's how:

df <- df |>
  mutate(new_date = as.Date(strptime(published_date, format = "%B %d, %Y %A")))
  
The format string specifies:

%B - full month name (January)
%d - day of the month (14)
%Y - 4-digit year (2025)
%A - full weekday name (Tuesday)

This will create a proper Date column that you can use for analysis and plotting. The resulting date will be in the standard format "2025-01-14".

#count years, publications, etc. create a ggplot